### 2023-09-26 doc/_single_results.qmd
small enhancements to section

To assess the inference pipeline performance, we applied it to a simulated
dataset with known ground truth relative fitness values (See `[ref to SI]` for
details on simulation). @fig-02(A) shows the structure of the synthetic dataset.
The majority of barcodes of interest (faint color lines) are adaptive compared
to the neutral barcodes ($s^{(m)} > 0$). This experimental condition is required
to reliably measure the relative fitness since all barcodes of interest occupy a
small fraction of the population at the beginning of the experiment `[ref SI
section on detrimental mutants]`. Although the barcode frequency trajectories
look relatively smooth, our fitness model requires the computation of the log
frequency ratio between adjacent time points as derived in @eq-logfreq.
@fig-02(B) shows such data transformation where we can better appreciate the
observational noise input into our statistical model. This noise is evident for
the darker lines representing the neutral barcodes since all of these lineages
are assumed to be identically distributed under this log frequency ratio
transformation.

To visualize the performance of our inference pipeline in fitting our fitness
model to the observed data, we compute the so-called posterior predictive checks
(PPC). In short, the PPC consists of repeatedly generating synthetic datasets in
agreement with the results from the inference results. In other words, we use
the resulting parameter values from the ADVI inference to generate possible
datasets in agreement with the inferred values. @fig-02(C) shows these results
for all neutral lineages (upper left corner plot) and a few representative
non-neutral barcodes. The different color shades represent the 95\%, 68\%, and
5\% credible regions, i.e., the regions where we expect to find the data with
the corresponding probability---or in terms of our parameter, the $X\%$ credible
region is the interval where we expect the true parameter to lie with $X\%$
probability.

@@

![**Single dataset inference**. (A) Frequency trajectories that represent the
raw data going into the inference. (B) Log frequency ratio between two adjacent
time points used by the inference pipeline. Darker lines represent the neutral
barcodes. These transformed data are much more noisy than the seemingly smooth
frequency trajectories. (B) Examples of the posterior predictive checks for all
neutral lineages (upper left panel) and a subset of representative mutant
lineages. Shaded regions represent the 95\%, 68\%, and 5\% credible regions for
the data. The reported errors above the plot represent the 68\% credible region
on the mutant relative fitness marginal distribution. (C) Comparison between the
ground truth fitness value from the logistic-growth simulation and the inferred
fitness value. Gray error bars represent the 68\% posterior credible region. (D)
The empirical cumulative distribution function (ECDF) for the absolute z-score
value of the ground truth parameter value within the inferred fitness posterior
distribution.](./figs/fig02){#fig-02}

>>>>

To assess the inference pipeline performance, we applied it to a simulated
dataset with known ground truth relative fitness values (See @sec-logistic for
details on simulation). @fig-02(A) shows the structure of the synthetic dataset.
The majority of barcodes of interest (faint color lines) are adaptive compared
to the neutral barcodes ($s^{(m)} > 0$). Although the barcode frequency
trajectories look relatively smooth, our fitness model requires the computation
of the log frequency ratio between adjacent time points as derived in
@eq-logfreq. @fig-02(B) shows such data transformation where we can better
appreciate the observational noise input into our statistical model. This noise
is evident for the darker lines representing the neutral barcodes since all of
these lineages are assumed to be identically distributed.

To visualize the performance of our inference pipeline in fitting our fitness
model to the observed data, we compute the so-called posterior predictive checks
(PPC). In short, the PPC consists of repeatedly generating synthetic datasets in
agreement with the results from the inference results. In other words, we use
the resulting parameter values from the ADVI inference to generate possible
datasets in agreement with the inferred values (See @sec-ppc for further details
on these computations). @fig-02(C) shows these results for all neutral lineages
(upper left corner plot) and a few representative non-neutral barcodes. The
different color shades represent the 95\%, 68\%, and 5\% credible regions, i.e.,
the regions where we expect to find the data with the corresponding
probability---or in terms of our parameter, the $X\%$ credible region is the
interval where we expect the true parameter to lie with $X\%$ probability.

@@

![**Single dataset inference**. (A) Frequency trajectories that represent the
raw data going into the inference. (B) Log frequency ratio between two adjacent
time points used by the inference pipeline. Darker lines represent the neutral
barcodes. These transformed data are much more noisy than the seemingly smooth
frequency trajectories. (C) Examples of the posterior predictive checks for all
neutral lineages (upper left panel) and a subset of representative mutant
lineages. Shaded regions represent the 95\%, 68\%, and 5\% credible regions for
the data. The reported errors above the plot represent the 68\% credible region
on the mutant relative fitness marginal distribution. (D) Comparison between the
ground truth fitness value from the logistic-growth simulation and the inferred
fitness value. Gray error bars represent the 68\% posterior credible region for
the relative fitness values. (E) The empirical cumulative distribution function
(ECDF) for the absolute z-score value of the ground truth parameter value within
the inferred fitness posterior distribution.](./figs/fig02){#fig-02}
